AXON. A Selection of Greek Historical Inscriptions. A Database for Research and Teaching

### 2015-12-17 source/21/EAGLE2016_Full_Paper_revised.tex
paper 21 updated
first doc to tex correction done

%%%%%%%%
%TITLE:%
%%%%%%%%
\title{}
\author{}
\date{}
\begin{document}
AXON. A Selection of Greek Historical Inscriptions

A Database for Research and Teaching

\bigskip

Abstract

The AXON Project has developed a database of Greek historical inscriptions, from the birth of the~polis~in the Archaic
Age to 31 BC. Each entry is provided with the object’s description, a complete lemma, Greek text with critical
apparatus, Italian translation and commentary with keywords and indexes, and updated bibliography. New insights for
data-inclusion have been developed. The database supports enlargement and offers a high degree of searchability. Our
aim is to illustrate the structure, the contents and the solutions we have come up with in the development of the AXON
Project. We will also offer some suggestions for teaching and academic research purposes.

Keywords

Online epigraphic editions; interoperability of digital editions of Greek historical inscriptions; images of Greek
historical inscriptions; digital epigraphy in teaching and research

\bigskip

1. The AXON Project

@@

AXON includes texts from the birth of the Greek polis in the Archaic Age\footnote{ See the Introduction in Hansen \&
Nielsen 2004; Hansen 2006. } to 31 BC, a chronological frame traditionally related to Greek History (though a future
extension of this chronological limit is not excluded). The epigraphic entries have been prearranged in order to allow
a wide and well-structured description of each document. At the same time, a common and coherent lexicon has been
produced, which will permit an easier indexing of significant words and will make future searches much quicker.

\bigskip

2. A unique model-entry for a great diversity of inscriptions: taxonomy and categorisation

\bigskip

2.1. Entry description

\bigskip

2.1.1. Object’s description

@@

\begin{enumerate}
\item Object type
\item Material
\item Object’s dimensions
\item State of preservation
\item Further descriptive elements
\item Date and context of finding
\item Finding site (modern nation, ancient region, ancient and modern name of the city, if known)
\item Actual location (modern nation, city, museum/archaelogical context, inventory number)
\end{enumerate}
The great majority of these categories can be selected from a given number of options from a pull-down menu. Some
categories – such as Object type, Material, or State of preservation – are directly linked with the corresponding
sections in the EAGLE Vocabularies (see http://www.eagle-network.eu/resources/vocabularies/). Furthermore, a hyperlink
has been created between the AXON-entries and Pleiades website (http://pleiades.stoa.org): where the finding site is
known, each entry offers the geographic coordinates and a Googlemaps visualisation. This gives the possibility of
rapidly gathering the information for any single ancient location and allows for searches directly from an interactive
map.

\bigskip

2.1.2. Chronology

The chronological delimitation of each text is supported by many options, as you can see in fig. 1:

\begin{figure}
\centering
\includegraphics[width=10.88cm,height=4.156cm]{EAGLE2016FullPaperrevised-img001.png}
\end{figure}
Fig. 1. Window for the input of data, section Text/Chronology

\bigskip

2.1.3. Alphabet \& language

\bigskip

@@

\begin{enumerate}
\item Type of Inscription (with link to EAGLE’s Vocabularies:
http://www.eagle-network.eu/resources/vocabularies/typeins/. The categories are those used by Guarducci 1967-1978)
\item Text’s structure
\item Writing (Execution technique\footnote{ With link to http://www.eagle-network.eu/resources/vocabularies/writing/.};
different types of epichoric alphabets according to Kirchhoff’s colour-coded map; Local script\footnote{ Following the
categorisation in Jeffery 1990.}; Palaeographic features and letters’ form\footnote{ The letter-form and glyph-form are
based on the symbols of the font Cardo (http://scholarsfonts.net/cardofnt.html), but many have been developed by the
AXON Team on the examples of letter-form given in Jeffrey 1990 (see also
http://poinikastas.csad.ox.ac.uk/browseGlyphs.shtml).}; letters’ heights, description and layout of the text field;
Direction of Text)
\item Language (with an option for any dialect’s peculiarities)
\end{enumerate}

\bigskip

\bigskip

2.1.4. Genetic lemma \& apparatus criticus

The text of each inscription is preceded by a hierarchically arranged lemma (the so called genetic lemma, according to
Louis Robert’s definition\footnote{ Robert \& Robert 1954.}) and is followed by the apparatus criticus.

\bigskip

2.1.5. Italian translation \& commentary

@@

\bigskip

2.1.6. Abstract

@@

\begin{figure}
\centering
\includegraphics[width=10.88cm,height=5.449cm]{EAGLE2016FullPaperrevised-img002.png}
\end{figure}
Fig. 2. Example of Abstract in AXON. The different colours allow a selection of words from a drop-down menu.

\bigskip

@@

\bigskip

2.1.7. Bibliography

Finally, an updated bibliography highlights any previous edition for each entry, as well as all the appropriate
secondary sources. SEG abbreviations have been used for epigraphic corpora and other publications (the section
“materiali” on the website gives access to a list of all abbreviations, useful for students, as well\footnote{
http://virgo.unive.it/venicepigraphy/axon/public/axon/pagine/materiali (still being processed).}).

\bigskip

2.2. Internal \& external interoperability of the AXON database

\bigskip

@@

\bigskip

2.3. A simple website interface for the input of data

\bigskip

@@

\bigskip

\bigskip

3. Searchability

\bigskip

The website is designed to allow for many search options. Beyond the “full text” search and another based on the number,
title and author of the entry (see fig. 3), three other search-possibilities will also be available:

\begin{enumerate}
\item browse all the entries according to the inscriptions’ a) typology, b) chronology, and c) area of origin;
\item access the entries through an interactive map;
\end{enumerate}

\begin{figure}
\centering
\includegraphics[width=10.88cm,height=7.354cm]{EAGLE2016FullPaperrevised-img003.png}
\end{figure}
Fig. 3. Search based on “Full text”, number, title, and author of the entry.

\bigskip

\begin{enumerate}
\item perform an advanced search based on different categories: 
\end{enumerate}
\begin{enumerate}
\item \begin{enumerate}
\item bibliography
\item keywords
\item object’s description and preservation (see fig. 4)
\item chronology
\item text (single words or phrases, typology, dialect, alphabet, letter-form, etc.) (see fig. 5).
\end{enumerate}
\end{enumerate}

\bigskip

\bigskip

\begin{figure}
\centering
\includegraphics[width=10.88cm,height=9.518cm]{EAGLE2016FullPaperrevised-img004.png}
\end{figure}
Fig. 4. Different searching categories.

\bigskip

\bigskip

\bigskip

\bigskip

\bigskip

\bigskip

Fig. 5. Advanced text search.

\bigskip

\begin{figure}
\centering
\includegraphics[width=10.88cm,height=9.888cm]{EAGLE2016FullPaperrevised-img005.png}
\end{figure}
Other filters will be employed for each search result. A section entitled “tools” will also be available, and will
include information on the entries’ structure, tables with the contents of different categories, links to Vocabularies
and websites, etc.

4. The AXON Project for teaching and academic research

\bigskip

4.1. Teaching

\bigskip

@@

\bigskip

\bigskip

4.2. Significance for the academic community

\bigskip

>>>>

%%%%%%%%
%TITLE:%
%%%%%%%%
\title{AXON. A Selection of Greek Historical Inscriptions. A Database for Research and Teaching}
 \author[cafo]{Ivan Matijašić\corref{first}}
\author[cafo]{Stefania De Vido}
\author[cafo]{Silvia Palazzo}
\address[cafo]{Dipartimento di Studi Umanistici, Università Ca' Foscari Venezia}
\cortext[first]{Corresponding author. Email: ivan.matijasic@unive.it}
\begin{document}

\maketitle
\begin{abstract}
The AXON Project has developed a database of Greek historical inscriptions, from the birth of the~polis~in the Archaic
Age to 31 BC. Each entry is provided with the object’s description, a complete lemma, Greek text with critical
apparatus, Italian translation and commentary with keywords and indexes, and updated bibliography. New insights for
data-inclusion have been developed. The database supports enlargement and offers a high degree of searchability. Our
aim is to illustrate the structure, the contents and the solutions we have come up with in the development of the AXON
Project. We will also offer some suggestions for teaching and academic research purposes.

\end{abstract}
\keywords{Online epigraphic editions; interoperability of digital editions of Greek historical inscriptions; images of Greek
historical inscriptions; digital epigraphy in teaching and research}

\section{The AXON Project }

@@

AXON includes texts from the birth of the Greek polis in the Archaic Age\footnote{ See the Introduction in \citet{nielsen_inventory_2004}; \citet{hansen_polis:_2006}. } to 31 BC, a chronological frame traditionally related to Greek History (though a future
extension of this chronological limit is not excluded). The epigraphic entries have been prearranged in order to allow
a wide and well-structured description of each document. At the same time, a common and coherent lexicon has been
produced, which will permit an easier indexing of significant words and will make future searches much quicker.

\section{A unique model-entry for a great diversity of inscriptions: taxonomy and categorisation}

\subsection{Entry description}

\subsubsection{Object’s description}

@@

\begin{enumerate}
\item Object type
\item Material
\item Object’s dimensions
\item State of preservation
\item Further descriptive elements
\item Date and context of finding
\item Finding site (modern nation, ancient region, ancient and modern name of the city, if known)
\item Actual location (modern nation, city, museum/archaelogical context, inventory number)
\end{enumerate}
The great majority of these categories can be selected from a given number of options from a pull-down menu. Some
categories – such as Object type, Material, or State of preservation – are directly linked with the corresponding
sections in the EAGLE Vocabularies (see \url{http://www.eagle-network.eu/resources/vocabularies/}). Furthermore, a hyperlink
has been created between the AXON-entries and Pleiades website (\url{http://pleiades.stoa.org}): where the finding site is
known, each entry offers the geographic coordinates and a Googlemaps visualisation. This gives the possibility of
rapidly gathering the information for any single ancient location and allows for searches directly from an interactive
map.

\subsubsection{Chronology}

The chronological delimitation of each text is supported by many options, as you can see in Fig.~\ref{fig:1}:

\begin{figure}[!bp]
\centering
 \includegraphics[width=\columnwidth]{EAGLE2016FullPaperrevised-img001.png}
\caption{Window for the input of data, section Text/Chronology}
\label{fig:1}
\end{figure}

\subsubsection{Alphabet \& language}

@@

\begin{enumerate}
\item Type of Inscription (with link to EAGLE’s Vocabularies:
\url{http://www.eagle-network.eu/resources/vocabularies/typeins/}. The categories are those used by \citet{guarducci_epigrafia_1987})
\item Text’s structure
\item Writing (Execution technique\footnote{ With link to \url{http://www.eagle-network.eu/resources/vocabularies/writing/}.};
different types of epichoric alphabets according to Kirchhoff’s colour-coded map; Local script\footnote{ Following the
categorisation in \citet{jeffery_local_1990}.}; Palaeographic features and letters’ form\footnote{ The letter-form and glyph-form are
based on the symbols of the font Cardo (\url{http://scholarsfonts.net/cardofnt.html}), but many have been developed by the
AXON Team on the examples of letter-form given in Jeffrey 1990 (see also
\url{http://poinikastas.csad.ox.ac.uk/browseGlyphs.shtml}).}; letters’ heights, description and layout of the text field;
Direction of Text)
\item Language (with an option for any dialect’s peculiarities)
\end{enumerate}

\subsubsection{Genetic lemma \& apparatus criticus}

The text of each inscription is preceded by a hierarchically arranged lemma (the so called genetic lemma, according to
Louis Robert’s definition\footnote{\citet{robert_carie:_1954}.}) and is followed by the apparatus criticus.

\subsubsection{Italian translation \& commentary}

@@

\subsubsection{Abstract}

@@

\begin{figure}[!bp]
\centering
 \includegraphics[width=\columnwidth]{EAGLE2016FullPaperrevised-img002.png}
\caption{Example of Abstract in AXON. The different colours allow a selection of words from a drop-down menu. }
\label{fig:2}
\end{figure}

@@

\subsubsection{Bibliography}

Finally, an updated bibliography highlights any previous edition for each entry, as well as all the appropriate
secondary sources. SEG abbreviations have been used for epigraphic corpora and other publications (the section
``materiali'' on the website gives access to a list of all abbreviations, useful for students, as well\footnote{
\url{http://virgo.unive.it/venicepigraphy/axon/public/axon/pagine/materiali} (still being processed).}).

\subsection{Internal \& external interoperability of the AXON database}

@@

\subsection{A simple website interface for the input of data}

@@

\section{Searchability}

The website is designed to allow for many search options. Beyond the ``full text'' search and another based on the number,
title and author of the entry (see Fig.~\ref{fig:3}), three other search-possibilities will also be available:

\begin{figure}[!bp]
\centering
 \includegraphics[width=\columnwidth]{EAGLE2016FullPaperrevised-img003.png}
\caption{Search based on ``Full text'', number, title, and author of the entry.}
\label{fig:3}
\end{figure}

\begin{enumerate}
\item browse all the entries according to the inscriptions’ a) typology, b) chronology, and c) area of origin;
\item access the entries through an interactive map;
\item perform an advanced search based on different categories: 
\begin{enumerate}
\item bibliography
\item keywords
\item object’s description and preservation (see Fig.~\ref{fig:4})
\item chronology
\item text (single words or phrases, typology, dialect, alphabet, letter-form, etc.) (see Fig.~\ref{fig:5}).
\end{enumerate}
\end{enumerate}

\begin{figure}[!bp]
\centering
 \includegraphics[width=\columnwidth]{EAGLE2016FullPaperrevised-img004.png}
\caption{Different searching categories.}
\label{fig:4}
\end{figure}

\begin{figure}[!bp]
\centering
 \includegraphics[width=\columnwidth]{EAGLE2016FullPaperrevised-img005.png}
\caption{Advanced text search}
\label{fig:5}
\end{figure}

Other filters will be employed for each search result. A section entitled ``tools'' will also be available, and will
include information on the entries’ structure, tables with the contents of different categories, links to Vocabularies
and websites, etc.

\section{The AXON Project for teaching and academic research}

\subsection{Teaching}

@@

\subsection{Significance for the academic community}
